Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-02-19 Paper.Rmd
finished with a rount of editing the Q3 sections, except still need to do the TL sup mat section

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing the potential for rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewinton-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of populations, the $i^{th}$ component $P_i(t)$ representing
the population in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed between-patch
Spearman correlations X and Pearson correlations Y in all simulations, 
<!--***DAN: Please fill in X and Y. If it was the Kendall that was used, please 
change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
If you don't know how to answer these questions, let me know and remind me where the 
code is that you used to generate the noise. Actually, even if you can do this yourself,
please insert a comment below this one indcating where the code is for model and noise
***Shya: I don't know the answer, the code which generates extreme right or left tail dep noise 
is written by you : ExtremeTailDep.R; there is no mention to preserve spearman or kendall or pearson.
I called retd function from ExtremeTaildep.R in another script mydispersal.R line 116, 128 
to generate right and left tail dep. noise.-->
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with populations dispersing either to nearest neighbors only,
or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
rate was $d$. After each time step, if the population in a patch was below $1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

Skewness results were generally consistent with hypotheses. For the datasets that
exhibited strong lower-tail (respectively, upper-tail) 
dependence in earlier analyses (namely, the green spruce 
aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
for upper-tail, Table 
\ref{tab_multivar_summary}), the skewness, through time, of the spatial average of the 
empirical data was less than (respectively, greater than) a significant fraction (i.e.,
more than $95\%$) of the
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For the datasets that demonstrated a 
more moderate lower-tail dependence in earlier analyses (the 
\emph{Ceratium furca} abundance and the methane data), skewness of the spatial
average showed a non-significant tendency toward being less than surrogate skewnesses
(Fig. \ref{fig_skewness_spearman}C), or was less than a marginally significant 
fraction of surrogate skewnesses (Fig. \ref{fig_skewness_spearman}D).

<!--***DAN: Dan has not edited the results for the extinction risk stuff yet-->
 We found, as we
hypothesized, left tail dependent environmental fluctuations increased extinction risk of metapopulations than the right-tail
dependent ones (Fig. \ref{fig_extrisk_lcohen_local}). When number of patches increased from $5$ to $25$, extinction risk 
decreased in both cases. We repeated our simulations using same parameters with global dispersal among patches 
and got similar results (see Fig. \ref{SM-fig_extrisk_lcohen_global}).

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
among two nearest neighbors ($r=0$, $d=$ rate of dispersal). 
Left-tail dependent environmental noise caused higher 
extinction risk than right-tail dependent one for number of patches $(A)$ $N = 5$ and $(B)$ $N = 25$, respectively,
where both case showed lower extinction risk with increasing number of patches.\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}
<!--***DAN: Shya, can you please change this figure as follows:
1) Make it only go to d=0.5. I think reviewers will complain that d close to 1 is a weird and 
unrealistic. Anyway, results are the same for larger d, so we are not hiding anything
2) the main comparison is between right and left tail dependence, so make the left panel show 
the numlocs=5 case, with results for both right- and left-tail dependence of the noise shown. 
And make the right panel show the numlocs=25 case with results for both right- and left-tail
dependence of the noise shown.
3) Right now you use grey dots on one panel and black on the other, but you can just use black 
on both.
4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
***Shya: Done.-->

>>>>

We also examined whether the nature of the
tail dependence of an environmental variable across space could influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. 
We provided a test of the conceptual validity of this hypothesis using a metapopulation
extension of the Lewinton-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{th}$ component $P_i(t)$ representing
the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
distributed in each patch, were independent through time, and showed the same 
spatial correlations (to within sampling error) for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
to be arranged in a line, with dispersal occurring either to nearest neighbors only,
or globally, in different simulations. The dispersal 
rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
if the population density in a patch was below $1$ (an extinction threshold), it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from initial populations in each patch of X, 
<!--***DAN: Shya, please fill in X-->
and calculated extinction risk after $25$ time steps.

@@

\noindent Skewness results were generally consistent with hypotheses. For the datasets that
exhibited strong lower-tail (respectively, upper-tail) 
dependence in earlier analyses (namely, the green spruce 
aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
for upper-tail, Table 
\ref{tab_multivar_summary}), the skewness, through time, of the spatial average of the 
empirical data was less than (respectively, greater than) a significant fraction (i.e.,
more than $95\%$) of the
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For the datasets that demonstrated a 
more moderate lower-tail dependence in earlier analyses (the 
\emph{Ceratium furca} abundance and the methane data), skewness of the spatial
average showed a non-significant tendency toward being less than surrogate skewnesses
(Fig. \ref{fig_skewness_spearman}C), or was less than a marginally significant 
fraction of surrogate skewnesses (Fig. \ref{fig_skewness_spearman}D).

As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
extinction risk for the spatially 
extended Lewinton-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
dispersal. We also examined the influence of non-N copula structure and tail dependence
on Taylor's law in Appendix \ref{SM-Consequence_TL_simulation}.

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for the metapopulation extension of the Lewinton-Cohen model, after 
$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}

### 2019-02-19 SupportingInformation.Rmd
finished with a rount of editing the Q3 sections, except still need to do the TL sup mat section

<!-- Taylor's law section moved to Paper.Rmd -->

@@

\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

@@

<!--How noise was generated ? code : ExtremeTailDep.R-->

<!--How dispersal matrix was generated ? code : mydispersal.R-->
A $N \times N$ dispersal matrix ($D$) was generated assuming that the $N$ number of metapopulation patches
were arranged in a linear chain fashion. Say, population from $i^{th}$ patch could either disperse to its nearest
connecting patches only (we called it as local dispersal) or could disperse to all other patches 
equally (we called it as global dispersal) at an dispersal rate, $d$. In case of 
local dispersal matrix the diagonal entries $D_{ii}$ would be equal to $d$ and its immediate neighbors
$D_{ii\pm j ,\;\ j=1}$ would get an equal proportion from the rest of $(1-d)$. The other distant neighbors 
$D_{ii\pm j ,\; j>1}$ were set to zero to execute local dispersal. For global dispersal matrix, $D_{ii}$
would be $d$ and all the other entries were equal to $(1-d)/N$. Code to generate these dispersal matrices 
was implemented in [Dispersal_mat](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R) function.

>>>>

@@

\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

@@

<!--How noise was generated ? code : ExtremeTailDep.R-->
\noindent Noise was generated by first producing a 25 by $N$ matrix, $m$, of 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each location. If left-tail dependent noise is desired, take the negative of the
entire matrix.

<!--How dispersal matrix was generated ? code : mydispersal.R-->
For local dispersal, the $N \times N$ dispersal matrix $D$ was generated by assuming, 
for simplicity, that the $N$ habitat patches
were arranged, evenly spaced, in a line. A fraction $d$ of each population dispersed during
any given time step, equally distributed to the two or one nearest neighbors. 
For global dispersal, the fraction $d$ that dispersed from each patch was divided up equally
among all other patches.
